Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-05-28 manuscript/manuscript.Rmd
updates Table 1 with 90th percentile distance

#IPM
tmpipm <- ddply(ipm_onestep, .(quad, t1, species), summarise,
                mean_pred = mean(cover.t1),
                var_pred = var(cover.t1),
                mean_obs = mean(obs.cover.t1),
                resid = mean(resid))
ipmstats <- ddply(tmpipm, .(species), summarise,
                rho = cor(mean_pred, mean_obs),
                mae = mean(abs(resid)),
                var = mean(var_pred),
                mean_cover = mean(mean_obs*100))
ipmstats$model <- "IPM"

#QBM
tmpqbm <- ddply(qbm_onestep, .(quad, year, species), summarise,
                mean_pred = mean(predcov),
                var_pred = var(predcov),
                mean_obs = mean(obscov),
                resid = mean(resid))
qbmstats <- ddply(tmpqbm, .(species), summarise,
                rho = cor(mean_pred, mean_obs),
                mae = mean(abs(resid)),
                var = mean(var_pred),
                mean_cover = mean(mean_obs*100))
qbmstats$model <- "QBM"

@@

stats_table <- rbind(ipmstats, qbmstats)
stats_table <- stats_table[ order(stats_table[,"species"]), ]
stats_table <- stats_table[,c("species", "model", "mae",
                              "rho","mean_cover")]
# comment          <- list()
# comment$pos      <- list()
# comment$pos[[1]] <- c(nrow(stats_table))
# comment$command  <- c(paste("\\hline \n",  # we`ll replace all default hlines with this and the ones below
#                             "The IPM MAE is significantly lower for HECO ($p = 3.3\times10^{-10}$) and POSE ($p = 0.00013$. MAEs are statisticially similar between models for BOGR and PASM.   \n",
#                             sep = ""))

colnames(stats_table) <- c("Species", "Model", "MAE", "Accuracy", "Mean Obs. Cover")
print.xtable(xtable(stats_table, caption = "Mean absolute error (MAE) and accuracy (Pearson's $\\rho$) for one-step-ahead forecast from both model types. Forecasts were made without random year effects; only climate covariates could explain year-to-year variation."),type="latex", comment=FALSE,
             include.rownames=FALSE, caption.placement="top")
```
The IPM MAE is significantly lower for HECO ($p = 3.3 \times 10^{-10}$) and POSE ($p = 0.00013$). MAEs are statisticially similar between models for BOGR and PASM.

>>>>

#IPM
tmpipm <- ddply(ipm_onestep, .(quad, t1, species), summarise,
                mean_pred = mean(cover.t1),
                var_pred = var(cover.t1),
                mean_obs = mean(obs.cover.t1),
                resid = mean(resid),
                quant_dist = quantile(cover.t1*100, probs = 0.95) - quantile(cover.t1, probs = 0.05))
ipmstats <- ddply(tmpipm, .(species), summarise,
                rho = cor(mean_pred, mean_obs),
                mae = mean(abs(resid)),
                var = mean(var_pred),
                mean_cover = mean(mean_obs*100),
                quant_dist = mean(quant_dist))
ipmstats$model <- "IPM"

#QBM
tmpqbm <- ddply(qbm_onestep, .(quad, year, species), summarise,
                mean_pred = mean(predcov),
                var_pred = var(predcov),
                mean_obs = mean(obscov),
                resid = mean(resid),
                quant_dist = quantile(predcov*100, probs = 0.95) - quantile(predcov, probs = 0.05))
qbmstats <- ddply(tmpqbm, .(species), summarise,
                rho = cor(mean_pred, mean_obs),
                mae = mean(abs(resid)),
                var = mean(var_pred),
                mean_cover = mean(mean_obs*100),
                quant_dist = mean(quant_dist))
qbmstats$model <- "QBM"

@@

stats_table <- rbind(ipmstats, qbmstats)
stats_table <- stats_table[ order(stats_table[,"species"]), ]
stats_table <- stats_table[,c("species", "model", "mae",
                              "quant_dist","mean_cover")]
# comment          <- list()
# comment$pos      <- list()
# comment$pos[[1]] <- c(nrow(stats_table))
# comment$command  <- c(paste("\\hline \n",  # we`ll replace all default hlines with this and the ones below
#                             "The IPM MAE is significantly lower for HECO ($p = 3.3\times10^{-10}$) and POSE ($p = 0.00013$. MAEs are statisticially similar between models for BOGR and PASM.   \n",
#                             sep = ""))

colnames(stats_table) <- c("Species", "Model", "MAE", "90% Distance", "Mean Obs. Cover")
print.xtable(xtable(stats_table, caption = "Mean absolute error (MAE) and accuracy (Pearson's $\\rho$) for one-step-ahead forecast from both model types. Forecasts were made without random year effects; only climate covariates could explain year-to-year variation. 90 Distance refers to the average distance between the upper and lower 90 percentiles of the 100 predicted values for each quadrat-year combination."),type="latex", comment=FALSE,
             include.rownames=FALSE, caption.placement="top")
```
The IPM MAE is significantly lower for HECO ($p = 3.3 \times 10^{-10}$) and POSE ($p = 0.00013$). MAEs are statisticially similar between models for BOGR and PASM.

### 2015-05-28 manuscript/manuscript.tex
updates Table 1 with 90th percentile distance

\begin{table}[ht]
\centering
\caption{Mean absolute error (MAE) and accuracy (Pearson's $\rho$) for one-step-ahead forecast from both model types. Forecasts were made without random year effects; only climate covariates could explain year-to-year variation.} 
\begin{tabular}{llrrr}
  \hline
Species & Model & MAE & Accuracy & Mean Obs. Cover \\ 
  \hline
BOGR & IPM & 4.77 & 0.90 & 9.16 \\ 
  BOGR & QBM & 5.30 & 0.88 & 9.22 \\ 
  HECO & IPM & 0.60 & 0.80 & 1.22 \\ 
  HECO & QBM & 2.03 & 0.84 & 1.25 \\ 
  PASM & IPM & 0.18 & 0.65 & 0.40 \\ 
  PASM & QBM & 0.17 & 0.87 & 0.40 \\ 
  POSE & IPM & 0.78 & 0.61 & 1.23 \\ 
  POSE & QBM & 1.30 & 0.57 & 1.25 \\ 
   \hline
\end{tabular}
\end{table}

>>>>

\begin{table}[ht]
\centering
\caption{Mean absolute error (MAE) and accuracy (Pearson's $\rho$) for one-step-ahead forecast from both model types. Forecasts were made without random year effects; only climate covariates could explain year-to-year variation. 90 Distance refers to the average distance between the upper and lower 90 percentiles of the 100 predicted values for each quadrat-year combination.} 
\begin{tabular}{llrrr}
  \hline
Species & Model & MAE & 90\% Distance & Mean Obs. Cover \\ 
  \hline
BOGR & IPM & 4.77 & 29.33 & 9.16 \\ 
  BOGR & QBM & 5.30 & 44.63 & 9.22 \\ 
  HECO & IPM & 0.60 & 2.82 & 1.22 \\ 
  HECO & QBM & 2.03 & 12.05 & 1.25 \\ 
  PASM & IPM & 0.18 & 0.49 & 0.40 \\ 
  PASM & QBM & 0.17 & 1.34 & 0.40 \\ 
  POSE & IPM & 0.78 & 2.16 & 1.23 \\ 
  POSE & QBM & 1.30 & 7.04 & 1.25 \\ 
   \hline
\end{tabular}
\end{table}
